---
title: "Day 9 Interactice"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(palmerpenguins)
library(kableExtra)
library(patchwork)
library(reprex)
```

Starting from penguins, create a new subset in which you:

1.  Only keep columns species, island, sex, body_mass_g, and flipper_length_mm

2.  Filter to only include Chinstrap and Gentoos

3.  Rename the species column to penguin_species

4.  Create a summary table that contains the mean of body mass, the mean of flipper length, and the total count of observations each by species and sex

```{r}
penguins_new <- penguins %>% 
  select(species,sex,body_mass_g,flipper_length_mm) %>% 
  filter(species %in% c("Chinstrap","Gentoo")) %>% 
  rename(penguin_species = species) %>% 
  group_by(penguin_species, sex) %>% 
  summarize(mean_body_mass = mean(body_mass_g, na.rm = TRUE),
            mean_flipper_length = mean(flipper_length_mm, na.rm = TRUE)
            ,total_count = n())


```

## A few variations on dplyr::across()

### Create a summary table of mean values for all numeric variables in the penguin dataset

```{r}
#One way that is tedious 
# penguins %>% 
#   group_by(species) %>% 
#   summarize(mean_mass = ...,
#             mean_flipper = ...,
#             # etc)


## Max values
penguins %>% 
  group_by(species) %>% 
  summarize(across(where(is.numeric), max, na.rm = TRUE))

```

## Calculate the mean values of all columns that start with "bill"

```{r}
penguins %>% 
  group_by(species,island) %>% 
  summarize(across(starts_with("bill"),mean, na.rm = TRUE))
```

### Write code that will return the minimum value of any columns that end with the character "mm", grouped by island and year

```{r}
penguins %>% 
  group_by(island,year) %>% 
  summarize(across(ends_with("mm"),min, na.rm = TRUE))
```

### Find and return both the mean and standard deviation of all columns starting with "bill", grouped by penguin species, and returning a new name that combines the original column with the function

```{r}
penguins %>% 
  group_by(species) %>% ## APPLY DIFFERENT FUNCTIONS AND ASSIGN DIFFERENT NAMES
  summarize(across(starts_with("bill"),list(mean = mean,
                                            sd = sd,
                                            max = max),
                   na.rm = TRUE,
                   .names = "{.col}_HI_{fn}"))



penguins %>% 
  group_by(species) %>% ## APPLY DIFFERENT FUNCTIONS AND ASSIGN DIFFERENT NAMES
  summarize(across(where(is.numeric),list(mean = mean,
                                            sd = sd,
                                            max = max),
                   na.rm = TRUE,
                   .names = "{.col}_HI_{fn}"))

```

```{r}

### How do we make this table look nicer?
penguins %>% 
  group_by(species, sex) %>%
  summarize(mean_mass = mean(body_mass_g, na.rm = TRUE),
            sd_mass = sd(body_mass_g, na.rm = TRUE)) %>% 
  kable(col.names = c("Species","Sex","Mean body mass (g)","SD body mass (g)")) %>% kable_styling(full_width = FALSE)
```

```{r}
chart1 <- penguins %>% 
  ggplot(aes(body_mass_g,flipper_length_mm)) + geom_point()

chart2 <- penguins %>% 
  ggplot(aes(bill_length_mm)) + geom_histogram()

## combine two charts using patchwork
chart1 + chart2

(chart1 + chart2 + chart1)*chart2
```

\

```{r}

#%>% fig-cap: Penguin body mass observations by species.
#%>% fig-alt: Penguin Jitterplot showing the body mass measurements (grams) for three pengun species (Adelie, Chinstrap, and Gentoo) observed near Palmer Archipelago, Antartica. Adelie and Chinstraps have similar body mass distributions, with Gentoos noticeable larger than both.


ggplot(penguins, aes(x = species, y = body_mass_g)) +
  geom_jitter(width = 0.2, alpha =0.5, color = 'navy')
```

```{r}

ggplot(data = mtcars, aes(x = mpg, y = hp)) + geom_point()


```
